Supplementary Material

### 2020-02-18 Revisions/Rebuttal.md

*This article presents 22 environmental drivers and uses these environmental drivers to identify areas of high cumulative exposure and characterize areas with similar cumulative exposure in the St. Lawrence System. In particular, the article introduces an open knowledge platform named eDrivers, though it doesn’t seem to be working at this moment (https://github.com/eDrivers - this site contains zero repositories, zero people, zero projects; the R package dDrivers is not available on CRAN). When this R package and the website become available, I believe the article and its related sources will be beneficial to the scientific and management communities. Despite its weaknesses (shown below), I believe this work is a good start for assessing and managing cumulative effects in the St. Lawrence System and can serve as a good example for other marine ecosystems.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
**Answer:**

> - The material has yet to be released publicly on GitHub as we were waiting for the review process to be completed. We can however provide access to the repositories in the interim.

> - As for the package, we will not be releasing a version on CRAN just yet. The material is not yet implemented on the SLGO servers (our partners on this project) and the data is currently contained directly in the package. Our goal is to publish the package and submit it for peer-review once the data is available and accessible through SLGO's API. However, the package will/can? be available on github in the interim.
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

@@

***Comment 1:***

*Authors need to make it clear up in the front that this work is only focusing on one of the four elements required to evaluate environmental impacts. Besides moving Lines 119-122 to around Line 77, it will be helpful to introduce the best practices in cumulative effects assessment and management: (1) selecting valued (ecosystem) components, (2) selecting indicators, (3) setting temporal and spatial scales, (4) addressing uncertainties, and (5) public participation. It will also be helpful to differentiate the two different indicators: effects-based and stressor-based. Effects-based indicators are direct measurements of the valued components (such as species abundance, biodiversity, etc.) and inherently encompass cumulative effects of multiple drivers. However, because of the poorly understood cause-effect linkages between stressors and valued components, it is currently hard to be used for decision-making. In comparison, stressor-based indicators (i.e., driver-based as this article defines and the type that this article uses) are usually better understood and can be proactively linked to management actions. However, these indicators do not capture the effects of all stressors and are not able to discern non-additive cumulative effects (e.g. synergism, antagonism). This point leads to the next weakness of this article.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Answer:***

> - We agree with reviewer 1 on the importance of clearly stating that we are solely focusing on drivers in our manuscript. The goal of our manuscript and the *eDrivers* initiative is to facilitate the widespread accessibility of such data for environmental assessments and integrative management. We believe it is clear that we are presenting but one element of a fully operational impact assessment given the title of the manuscript and the 3rd sentence in the abstract: "Our objective in this paper is to systematically characterize the distribution and intensity of drivers in the St. Lawrence System.". Our goal is  not to discuss cumulative effects assessments in particular. It is rather to provide knowledge on an element common to most holistic approaches, *i.e.* drivers characterization. We have strived to make this as clear as possible throughout the manuscript, namely in the introduction and in the perspectives sections, as noted by reviewer 1. As such, and considering the focus of the manuscript, we are hesitant to accept the suggestion of reviewer 1 to discuss cumulative effects assessment, best practices and indicator types at the beginning of the introduction. We feel that, considering the objective of this manuscript, the beginning of the introduction should rather focus on drivers specifically.

### 2020-02-26 Revisions/Beauchesne-eDriversMS-SI.rmd
include final revisions to the manuscript before sending to co-authors

---
title: Supplementary Material
fontsize: 12pt
output:
  pdf_document:
    toc: false
    number_sections: true
  html_document:
    number_sections: true
header-includes:
    - \usepackage{float}
    - \usepackage{setspace}
    - \onehalfspacing
relativeurls: true
bibliography: eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

<!--
cd phdproj/edrivers/edriversms
rmarkdown::render('./eDriversMS-SI-v2.rmd', 'pdf_document')
rmarkdown::render('./eDriversMS-SI-v2.rmd', 'word_document')
-->

@@

```{r tableS1, echo=F, fontsize=4}
load('./tables/tableS1.RData')
tableS1
```

@@

Oceans are the largest reservoir and sink of atmospheric carbon dioxide ($CO_2$).
Its uptake increases seawater acidity and the lowers saturation state with
respect to calcite ($\Omega_C$) and aragonite ($\Omega_A$), the two most common
$CaCO_3$ polymorphs constituting the shells and skeleton of many marine
organisms [@mucci2017]. This can have deleterious effects on carbonate-secreting
organisms (*e.g.* mollusks and crustaceans) and certain physiological processes
in non-calcifying organisms [@fabry2008; @kroeker2013].

@@

We used $\Omega_{A}$ saturation state to characterize ocean acidification in the
bottom waters of the St. Lawrence. The data come from the Department of
Fisheries and Oceans' (DFO) Atlantic Zone Monitoring Program [AZMP; @galbraith2018]
surveys and DFO's multispecies surveys [@bourdages2018] collected in
August-September of 2017 [@starr2019]. The carbonate chemistry was determined
through pH and TA measurements. Samples for pH and total alkalinity (TA) were
collected under a bubble free and no head space conditions into 500 $mL$
borosilicate glass flasks, 250 $\mu L$ of the saturated $HgCl_2$ solution was
added to sample, and processed following the "Guide to best practices for Ocean
CO2 Measurements" [@dickson2007].

$pH_T$ was determined spectrophotometrically using the indicator dye m-cresol
purple (Sigma-Aldrich). Absorbance was measured at 730, 578 and 434 nm before
and after dye addition in 10 cm quartz cells thermostated at 25 ± $0.05^oC$
[@dickson2007]. A similar procedure was carried out before each set of sample
measurements using a TRIS (Tris (hydroxymethyl) -aminomethane) buffer prepared
at a practical salinity (S) of approximately 30 [@millero1986]. Certified
Reference Material (CRM) (supplied by Professor Andrew Dickson, Scripps
Institution of Oceanography, San Diego, USA) was used for quality control of
our pH TRIS buffer.

TA was determined by potentiometric titration in an open cell using an
automated Radiometer potentiometric titrator (Titrilab 865) and a pH combination
electrode (pHC2001) in a continuous titrant addition mode, an algorithm
specifically designed for shallow end-point detection [@dickson2007]. The dilute
HCl titrant 0.1M in a solution of NaCl of 0.6M was calibrated using CRM provided
by Professor Andrew G. Dickson.

The carbonate system parameters (including $\Omega_{A}$) were
corrected for in situ pressure and temperature using the algorithm CO2SYS
[@lewis1998] with measured pH, total alkalinity, soluble reactive
and silicate concentrations as input parameters.

$\Omega_{A}$, pH, dissolved oxygen ($O_2$) and pressure were sampled at 117
stations in the Estuary and Gulf of St. Lawrence in the summer of 2018. The
majority of sampling stations were located in the Southern Gulf ($n = 84$),
compared to the Estuary and Northern Gulf ($n = 33$).

@@

There are likely different dynamics at play in the Southern Gulf, such as
waters with lowered alkalinity and salinity caused by the St. Lawrence plume.
Such water masses are have lower buffering capacities to decreasing pH and
$\Omega_{A}$. Also, the accumulation of metabolic carbon dioxide produced by
rapid oxic degradation of organic matter in the sandy
and permeable sediments of the Magdalen Shelf or through the St. Lawrence plume
could decrease $\Omega_{A}$ in the Southern Gulf (Starr and Chassé, *personal
communication*).

We thus divided the analyses between the Southern and Northern St. Lawrence.
The St. Lawrence was divided using the 350m isobath on the southern
slope of the Laurentian Channel, from the Cabot Strait to the tip of the Gaspé
Peninsula. $\Omega_A$ distribution in the Southern St. Lawrence was modelled
using the exponential kriging model.

We predicted $\Omega_A$ using $O_2$ for the Northern St. Lawrence. Whereas
there are few published datasets on $\Omega_{A}$ in the St.
Lawrence, there is a robust time-series on oxygen concentration in the water
column from the Department of Fisheries and Oceans’ (DFO) Atlantic Zone
Monitoring Program [AZMP; @galbraith2018]. We used oxygen concentration data
collected between 2013 and 2017
[@blais2019] and interpolated oxygen % saturation using cokriging with depth as a
covariable, as done in [@dutil2011]. Bathymetric data come from @dutil2011
and have a 100 $km^2$ resolution. For more details, refer to the hypoxia
description. The *Deep* model was used to transform $O_2$ values to $\Omega_A$
in grid cells with depth > 350m, while the *Shallow* model was
applied to cells with depth < 350m.

Based on the wide range of $\Omega_A$ levels at which organism are negatively
affected (see above), we built an index of acidification stress ($A_s$)
ranging between 0 and 1 that slowly begins increasing as $\Omega_A$ $\approx$
$2$, increases more rapidly close to $\Omega_A$ $= 1$ and reaches its peak at
$\Omega_A$ $\approx$ $0.5$:

@@

The data used to characterize hypoxia uses oxygen saturation in % and come from
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018] surveys in late spring and fall, as well as from annual
DFO’s multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. We provide a brief summary of data and methods to
describe hypoxia in this document. For more details, refer to @blais2019.

Oxygen concentration is measured at every station using an oxygen probe
(Sea-Bird SBE43) mounted on the CTD. The probe is calibrated against seawater
samples that are analyzed by Winkler titration on every cast
[see @blais2019 for calibration procedure]. The data used is the last depth
sampled on the CTD profile, which typically corresponds to ~10m above the
bottom. Oxygen concentration ($\mu M$ $L^{-1}$) is converted to oxygen
saturation taking into account the salinity and temperature at the selected
depth.

@@

We used temperature anomalies, *i.e.* deviations from long-term normal
conditions, to measure an annual index of stress associated with extreme
temperatures between 2013 and 2017. Temperature anomalies were calculated using
the difference between grid cell values with 1981-2010 climatological averages.
Anomaly time series were normalized by their standard deviation (SD) to allow
comparison between areas of the St. Lawrence with different temperature ranges.
For example, temperatures observed in deep channels are less variable than in
shallower regions of the St. Lawrence. Hence, if anomalies were expressed in
degrees Celsius, it would underestimate the relative importance of anomalies
in deep channels when compared to shallower regions. Grid cells whose monthly
value exceeded ±0.5 standard deviation (SD) from the long-term average were
considered as anomalous [@galbraith2018]. Outliers in the data were defined as
those that fell beyond the interquartile range * 3, identified as extreme
outliers by @tukey1977. Outlier values were capped to correspond to the 5th and
95th percentiles. Anomalies were divided into positive and negative anomalies
and the absolute value of anomalies were used as an indicator of the intensity
of bottom temperature anomalies. The mean anomaly intensity between 2013 and
2017 for each grid cell was used to generate the final index of bottom
temperature anomalies.

@@

The data used to characterize sea level rise risk comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

Aquaculture data comes from a variety of sources in the St. Lawrence because
aquaculture sites are mostly managed at the provincial level. We therefore had
to gather the data on aquaculture sites from the 5 provinces dividing the
St. Lawrence [@mapaq2016; @dfo2016a; @aaf2016; @fa2016; @ffa2016].

Invertebrates aquaculture is especially important in the southern and western
Gulf. Fish and algae aquaculture, on the other hand, remains marginal.
Considering this, we only considered invertebrates aquaculture for the
aquaculture driver layer. However, if fish or algae farming were to become more
important, these driver should be incorporated in future analyses as individual
layers, as impacts vary between types of aquaculture.

Aquaculture activities are highly localized and potential effects do not or
rarely extend beyond the location of the farms. We therefore only considered
the actual location of sites to characterize the distribution of this driver.
We were unable to characterize site production in terms of biomass farmed, which
could provide an indication of the intensity of aquaculture activities. As such,
we considered aquaculture as binary presence-absence data in our analyses.

@@

We used lights at night as a proxy of coastal infrastructure development, as
terrestrial stable lights at night represent light from human settlements and
industrial sites with electricity.

The data comes from the Nighttime Lights Time Series. Nighttime light products
are compiled by the Earth Observation Group at the National Oceanic and
Atmospheric Administration's (NOAA) National Centers for Environmental
Information (NCEI). They use globally available nighttime data obtained from
the Visible Infrared Imaging Radiometer Suite (VIIRS) Day/Night Band (DNB) of
the Defense Meteorological Satellite Program (DMSP) to characterize global
average radiance ($nanoWatts$ $cm^{-2}$ $sr^{-1}$) composite images at a
15 arc-second (~200 m) resolution.

@@

The data used to characterize inorganic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Inorganic pollution was modelled using impervious surface area (*i.e.*
artificial surfaces such as paved roads) under the assumption that most of this
pollution source comes from urban runoff. Inorganic pollution originating from
point-sources or in areas lacking paved roads is therefore not captured by
this layer. The data obtained was aggregated at the watershed scale and spread
into coastal and marine environments was modelled using a diffusive plume model
from each watershed pourpoints (*e.g.* river mouths).

@@

The data used to characterize nutrient pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual fertilizer use in tonnes ($t$) was used as a proxy of nutrient pollution.
The data used came from the Food and Agriculture Organization of the United
Nations (FAO). Gaps in data were modelled using a linear regression between
fertilizer and pesticides or agricultural gross domestic product (GDP).
Dasymetric maps were then used to distribute fertilizer data over the landscape
using 2009 data from the Moderate Resolution Imaging Spectroradiometer (MODIS)
at ~500 m resolution and aggregated to watersheds. Diffusive plume models
from each watershed pourpoint (*e.g.* river mouths) were then used to model the
distribution and intensity of nutrient pollution in coastal and marine
environments.

@@

The data used to characterize organic pollution comes from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual pesticide use in tonnes ($t$) was used as a proxy of organic pollution.
The data used came from the FAO and gaps in data were modelled using a linear
regression between pesticides and fertilizers or agricultural GDP. The same
methodology as that used to characterized nutrient pollution was then applied
to organic pollution.

@@

The data we use to describe the risk of toxic algae comes from an expert based
map delineating the areas where coastal areas are at risk from five different
toxins [@bates2019]. The map presents coastal areas at risk from 5
different toxins: 1) paralytic shellfish poisoning (PSP) toxins from the regular
presence of the dinoflagellate *Alexandrium catenella* (previously
*Alexandrium tamarense*) at high concentrations, 2) amnesic shellfish
poisoning (ASP) toxins from domoic acid 3) diarrhetic shellfish poisoning (DSP)
toxins, 4) spirolides and 5) pectenotoxins,  two toxins produced by
dinoflagellates occurring in the St. Lawrence.

@@

The impacts of fisheries activities in the St. Lawrence are evaluated using
DFO's fisheries logbooks program [@dfo2016]. While logbooks are not mandatory
for all fisheries in the St. Lawrence, they still provide a very thorough
overview of the spatial distribution and intensity of fishing activities in the
St. Lawrence. The data we used spans 6 years from 2010 to 2015 and details
218323 fishing events (36387 $\pm$ 3147 fishing events per year). There were 31
targeted species and a total of 53 caught species in the dataset.

Fishing activities are performed using a variety of gear types: trap, trawl,
dredge, driftnet, hand line, longline, scuba diving, purse seine, seine, beach
seine and jig fishing. Intensity of fishing activities was divided among gear
types and based on their respective types of environmental impacts (Table S2).
For example, traps and trawls have very different effects on a system. Gear
classification was done using the classification presented in @halpern2008a and
@halpern2015 and is broken down into 5 distinct classes: demersal destructive
(DD), demersal, non-destructive, low-bycatch (DNL), demersal, non-destructive,
high-bycatch (DNH), pelagic, low-bycatch (PLB) and pelagic, high-bycatch (PHB).
This categorization therefore divides the fisheries data into 5 distinct
driver layers characterizing fishing activities.

Gear types can also be further classified into fixed or mobile engines based
on their mobility. We used these two mobility classes to generate a buffer of
impact around each fishing activity coordinates to consider potential spatial
uncertainty associated with locations and the fact that mobile engines can be
tracted over several kilometers during fishing activities and that we do not
have the beginning and end points of mobile fishing events. Buffer sizes for
fixed and mobile engine was of 200 and 2000 meters, respectively.

@@

In order to characterize the intensity of fishing activities ($FI$), we used
a biomass yield density index. We multiplied the total annual biomass captured
in each grid cell *j*, regardless of species, by the proportion of fishing area
in each grid cell:

@@

The data used to characterize shipping comes from the global cumulative impacts
assessment on habitats [@halpern2008a; @halpern2015a] and available on the NCEAS
online data repository [@halpern2015]. We provide a brief summary of data and
methods in this document. For more details, refer to @halpern2015.

@@

The second set of data comes from the Automatic Identification System (AIS),
an initiative launched in 2002 that sought to improve marine safety by providing
mariners with real-time vessel traffic. Through the
International Maritime Organization SOLAS agreement, all vessels of over 300
gross tonnage on international voyages and those carrying passengers are now
required to be equipped with AIS transceivers. These transceivers use Global
Positioning System technology to locate vessels every 10 minutes. The data used
was from November 2010 to December 2011.

Data used come mostly from vessels that move globally (*i.e.* cargo, tanker and
passenger), as they are required to carry AIS transceivers, but also include
data from fishing, high-speed, pleasure and support classes. Shipping intensity
was evaluated as the number of fishing tracks at a 0.1 decimal degrees resolution.
For more details on data and methods used, consult [@walbridge2013].

>>>>

---
title: Supplementary Material
fontsize: 12pt
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
  html_document:
    number_sections: true
header-includes:
    - \usepackage{float}
    - \usepackage{setspace}
    - \onehalfspacing
    - \usepackage[svgnames]{xcolor}
    - \usepackage[normalem]{ulem}
    - \definecolor{blue}{HTML}{20cad5}
relativeurls: true
bibliography: Beauchesne-eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

<!--
rmarkdown::render('./Revisions/Beauchesne-eDriversMS-SI.rmd')
-->

@@

```{r tableS1, echo=F, fontsize=4}
load('../tables/tableS1.RData')
tableS1
```

@@

Oceans are the largest reservoir\textcolor{blue}{s} and sink\textcolor{blue}{s} of atmospheric carbon dioxide ($CO_2$).
Its uptake increases seawater acidity and \textcolor{red}{\sout{the}} lowers saturation state with
respect to calcite ($\Omega_C$) and aragonite ($\Omega_A$), the two most common
$CaCO_3$ polymorphs constituting the shells and skeleton\textcolor{blue}{s} of many marine
organisms [@mucci2017]. This can have deleterious effects on carbonate-secreting
organisms (*e.g.* mollusks and crustaceans) and certain physiological processes
in non-calcifying organisms [@fabry2008; @kroeker2013].

@@

We used $\Omega_{A}$ saturation state to characterize ocean acidification in the
bottom waters of the St. Lawrence. The data come from the Department of
Fisheries and Oceans' (DFO) Atlantic Zone Monitoring Program [AZMP; @galbraith2018]
surveys and DFO's multispecies surveys [@bourdages2018] collected in
August-September of 2017 [@starr2019].
\textcolor{blue}{$\Omega_{A}$, pH, dissolved oxygen ($O_2$) and pressure were sampled at 117
stations in the Estuary and Gulf of St. Lawrence in the summer of 2018. The
majority of sampling stations were located in the Southern Gulf ($n = 84$),
compared to the Estuary and Northern Gulf ($n = 33$).}

The carbonate chemistry was determined
through pH and \textcolor{blue}{total alkalinity} (TA) measurements. Samples for pH and \textcolor{red}{\sout{total alkalinity}} TA were
collected under \textcolor{red}{\sout{a}} bubble free and no head space conditions into 500 $mL$
borosilicate glass flasks. 250 $\mu L$ of \textcolor{red}{\sout{the}} saturated $HgCl_2$ solution was
added to sample, and processed following the "Guide to best practices for Ocean
CO2 Measurements" [@dickson2007].

$pH_T$ was determined spectrophotometrically using the indicator dye m-cresol
purple (Sigma-Aldrich). Absorbance was measured at 730, 578 and 434 nm before
and after dye addition in 10 cm quartz cells thermostated at 25 ± $0.05^oC$
[@dickson2007]. A similar procedure was carried out before each set of sample
measurement\textcolor{red}{\sout{s}} using a TRIS (Tris (hydroxymethyl) -aminomethane) buffer prepared
at a practical salinity (S) of approximately 30 [@millero1986]. Certified
Reference Material (CRM) (supplied by Professor Andrew Dickson, Scripps
Institution of Oceanography, San Diego, USA) was used for quality control of
our pH TRIS buffer.

TA was determined by potentiometric titration in an open cell using an
automated Radiometer potentiometric titrator (Titrilab 865) and a pH combination
electrode (pHC2001) in a continuous titrant addition mode, an algorithm
specifically designed for shallow end-point detection [@dickson2007]. The dilute\textcolor{blue}{d}
HCl titrant 0.1M in a solution of NaCl of 0.6M was calibrated using CRM provided
by Professor Andrew G. Dickson.

The carbonate system parameters (including $\Omega_{A}$) were
corrected for *in situ* pressure and temperature using the algorithm CO2SYS
[@lewis1998] with measured pH, \textcolor{red}{\sout{total alkalinity}} \textcolor{blue}{TA}, soluble reactive
and silicate concentrations as input parameters.

\textcolor{red}{\sout{$\Omega_{A}$, pH, dissolved oxygen ($O_2$) and pressure were sampled at 117
stations in the Estuary and Gulf of St. Lawrence in the summer of 2018. The
majority of sampling stations were located in the Southern Gulf ($n = 84$),
compared to the Estuary and Northern Gulf ($n = 33$).}}

@@

There are likely different dynamics at play in the Southern Gulf, such as
waters with lowered alkalinity and salinity caused by the St. Lawrence plume.
Such water masses \textcolor{red}{\sout{are}} have lower buffering capacities to decreasing pH and
$\Omega_{A}$. Also, the accumulation of metabolic carbon dioxide produced by
rapid oxic degradation of organic matter in the sandy
and permeable sediments of the Magdalen Shelf or through the St. Lawrence plume
could decrease $\Omega_{A}$ in the Southern Gulf (Starr and Chassé, *personal
communication*).

We thus divided the analyses between the Southern and Northern St. Lawrence.
The St. Lawrence was divided using the 350m isobath on the southern
slope of the Laurentian Channel, from the Cabot Strait to the tip of the Gaspé
Peninsula. $\Omega_A$ distribution in the Southern St. Lawrence was modeled
using the exponential kriging model.

We predicted $\Omega_A$ using $O_2$ for the Northern St. Lawrence. \textcolor{red}{\sout{Whereas}} \textcolor{blue}{While}
there are few published datasets on $\Omega_{A}$ in the St.
Lawrence, there is a robust time-series on oxygen concentration in the water
column from the Department of Fisheries and Oceans’ (DFO) Atlantic Zone
Monitoring Program [AZMP; @galbraith2018]. We used oxygen concentration data
collected between 2013 and 2017
[@blais2019] and interpolated oxygen % saturation using cokriging with depth as a
covariable, as done in [@dutil2011]. Bathymetric data come from @dutil2011
and have a 100 $km^2$ resolution. For more details, refer to the hypoxia \textcolor{blue}{data layer}
description. The *Deep* model was used to transform $O_2$ values to $\Omega_A$
in grid cells with depth > 350m, while the *Shallow* model was
applied to cells with depth < 350m.

Based on the wide range of $\Omega_A$ levels at which organism are negatively
affected (see above), we built an index of acidification stress ($A_s$)
ranging between 0 and 1 that slowly \textcolor{red}{\sout{begins increasing}} \textcolor{blue}{increases} as $\Omega_A$ $\approx$
$2$, increases more rapidly close to $\Omega_A$ $= 1$ and reaches its peak at
$\Omega_A$ $\approx$ $0.5$:

@@

The data used to characterize hypoxia \textcolor{red}{\sout{uses oxygen saturation in % and}} come from
the Department of Fisheries and Oceans’ (DFO) Atlantic Zone Monitoring Program
[AZMP; @galbraith2018] survey\textcolor{red}{\sout{s}} in late spring and fall, as well as from \textcolor{red}{\sout{annual}}
DFO’s \textcolor{blue}{annual} multispecies surveys for the northern Gulf in August and for the Magdalen
Shallows in September. We provide a brief summary of data and methods to
describe hypoxia in this document. For more details, refer to @blais2019.

Oxygen concentration is measured at every station using an oxygen probe
(Sea-Bird SBE43) mounted on \textcolor{red}{\sout{the}} \textcolor{blue}{a} CTD. The probe is calibrated against seawater
samples that are analyzed by Winkler titration on every cast
[see @blais2019 for calibration procedure]. The data used is the last depth
sampled on the CTD profile, which \textcolor{blue}{is} typically \textcolor{red}{\sout{corresponds to}} ~10m above the
bottom. Oxygen concentration ($\mu M$ $L^{-1}$) is converted to oxygen
saturation\textcolor{blue}{,} taking into account the salinity and temperature at the selected
depth.

@@

We used temperature anomalies, *i.e.* deviations from long-term normal
conditions, to measure an annual index of stress associated with extreme
temperatures between 2013 and 2017. Temperature anomalies were calculated using
the difference between grid cell values with 1981-2010 climatological averages.
Anomaly time series were normalized by their standard deviation (SD) to allow
comparison\textcolor{blue}{s} between areas of the St. Lawrence with different temperature ranges.
For example, temperatures observed in deep channels are less variable than in
shallower regions of the St. Lawrence. Hence, if anomalies were expressed in
degrees Celsius, it would underestimate the relative importance of anomalies
in deep channels when compared to shallower regions. Grid cells whose monthly
value exceeded ±0.5 standard deviation (SD) from the long-term average were
considered as anomalous [@galbraith2018]. Outliers in the data were defined as
those that fell beyond the interquartile range * 3, identified as extreme
outliers by @tukey1977. Outlier values were capped to correspond to the 5th and
95th percentile\textcolor{red}{\sout{s}} \textcolor{blue}{values}. Anomalies were divided into positive and negative anomalies
and the absolute value of anomalies were used as an indicator of the intensity
of bottom temperature anomalies. The mean anomaly intensity between 2013 and
2017 for each grid cell was used to generate the final index of bottom
temperature anomalies.

@@

The data used to characterize sea level rise risk come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

@@

\textcolor{red}{\sout{Aquaculture data comes from a variety of sources in the St. Lawrence because
aquaculture sites are mostly managed at the provincial level.
We therefore had to gather the data on aquaculture sites from the 5 provinces dividing the
St. Lawrence}}
\textcolor{blue}{As aquaculture sites are managed at the provincial level in the St. Lawrence, data on aquaculture sites come from various sources}
[@mapaq2016; @dfo2016a; @aaf2016; @fa2016; @ffa2016].
Invertebrates aquaculture is especially important in the southern and western
Gulf. Fish and algae aquaculture, on the other hand, remains marginal.
Considering this, we only \textcolor{red}{\sout{considered}} \textcolor{blue}{included} invertebrates aquaculture for the
aquaculture driver layer. However, if fish or algae farming were to become more
important, these driver\textcolor{blue}{s} should be incorporated in future analyses as individual
layers, as impacts vary between types of aquaculture.

Aquaculture activities are highly localized and potential effects do not or
rarely extend beyond the location of the farms. We therefore only considered
the \textcolor{red}{\sout{actual}} location of sites to characterize the distribution of this driver.
We were unable to characterize site production in terms of biomass farmed, which
could provide an indication of the intensity of aquaculture activities. As such,
we considered aquaculture as binary presence-absence data \textcolor{red}{\sout{in our analyses}}.

@@

\textcolor{red}{\sout{We used lights at night as a proxy of coastal infrastructure development, as
terrestrial stable lights at night represent light from human settlements and
industrial sites with electricity.}}
\textcolor{blue}{Terrestrial stable lights at night mostly represent light from human settlements and industrial sites with electricity.
We thus used lights at night as a proxy of coastal infrastructure development.}

The data come from the Nighttime Lights Time Series. Nighttime light products
are compiled by the Earth Observation Group at the National Oceanic and
Atmospheric Administration's (NOAA) National Centers for Environmental
Information (NCEI). They use globally available nighttime data obtained from
the Visible Infrared Imaging Radiometer Suite (VIIRS) Day/Night Band (DNB) of
the Defense Meteorological Satellite Program (DMSP) to characterize global
average radiance ($nanoWatts$ $cm^{-2}$ $sr^{-1}$) composite images at a
15 arc-second (~200 m) resolution.

@@

The data used to characterize inorganic pollution come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Inorganic pollution was modeled using impervious surface area (*i.e.*
artificial surfaces such as paved roads) under the assumption that most of this
pollution source comes from urban runoff. Inorganic pollution originating from
point-sources or in areas lacking paved roads is therefore not captured by
this layer. The data obtained was aggregated at the watershed scale. \textcolor{red}{\sout{and spread}}
\textcolor{blue}{Spread} into coastal and marine environments was modeled using a diffusive plume model
from each watershed pourpoints (*e.g.* river mouths).

@@

The data used to characterize nutrient pollution come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual fertilizer use in tonnes ($t$) was used as a proxy of nutrient pollution.
The data used came from the Food and Agriculture Organization of the United
Nations (FAO). Gaps in data were modeled using a linear regression between
fertilizer and pesticides or agricultural gross domestic product (GDP).
Dasymetric maps were then used to distribute fertilizer data over the landscape
using 2009 data from the Moderate Resolution Imaging Spectroradiometer (MODIS)
at ~500 m resolution and aggregated to watersheds. Diffusive plume models
from each watershed pourpoint (*e.g.* river mouths) were then used to model the
distribution and intensity of nutrient pollution in coastal and marine
environments.

@@

The data used to characterize organic pollution come from the global
cumulative impacts assessment on habitats [@halpern2008a; @halpern2015a] and
available on the NCEAS online data repository [@halpern2015]. We provide a
brief summary of data and methods in this document. For more details, refer to
@halpern2015.

Annual pesticide use in tonnes ($t$) was used as a proxy of organic pollution.
The data used came from the FAO and gaps in data were modeled using a linear
regression between pesticides and fertilizers or agricultural GDP. The same
methodology as that used to characterized nutrient pollution was then applied
to organic pollution.

@@

The data we use to describe the risk of toxic algae come from an expert based
map delineating the areas where coastal areas are at risk from five different
toxins [@bates2019]. The map presents coastal areas at risk from 5
different toxins: 1) paralytic shellfish poisoning (PSP) toxins from the regular
presence of the dinoflagellate *Alexandrium catenella* (previously
*Alexandrium tamarense*) at high concentrations, 2) amnesic shellfish
poisoning (ASP) toxins from domoic acid 3) diarrhetic shellfish poisoning (DSP)
toxins, 4) spirolides and 5) pectenotoxins, two toxins produced by
dinoflagellates occurring in the St. Lawrence.

@@

The impacts of fisheries activities in the St. Lawrence are evaluated using
DFO's fisheries logbooks program [@dfo2016]. While logbooks are not mandatory
for all fisheries in the St. Lawrence, they still provide a very thorough
overview of the spatial distribution and intensity of fishing activities \textcolor{red}{\sout{in the
St. Lawrence}}. The data we used span 6 years from 2010 to 2015 and detail
218323 fishing events (36387 $\pm$ 3147 fishing events per year). There were 31
targeted species and a total of 53 \textcolor{red}{\sout{caught}} species \textcolor{blue}{caught} in the dataset.

Fishing activities are performed using a variety of gear types: trap, trawl,
dredge, driftnet, hand line, longline, scuba diving, purse seine, seine, beach
seine and jig fishing. Intensity of fishing activities was divided among gear
types and based on their respective types of environmental impacts (Table S2).
For example, traps and trawls have very different effects on a system. Gear
classification was done using the classification presented in @halpern2008a and
@halpern2015 and is broken down into 5 distinct classes: demersal destructive
(DD), demersal, non-destructive, low-bycatch (DNL), demersal, non-destructive,
high-bycatch (DNH), pelagic, low-bycatch (PLB) and pelagic, high-bycatch (PHB).
This categorization \textcolor{red}{\sout{therefore}} divides the fisheries data into 5 distinct
driver layers characterizing fishing activities.

Gear types can \textcolor{red}{\sout{also}} be further classified into fixed or mobile engines based
on their mobility. We used these two mobility classes to generate a buffer of
impact around each fishing activity coordinates to consider potential spatial
uncertainty associated with locations, \textcolor{red}{\sout{and}} the fact that mobile engines can be
tracted over several kilometers during fishing activities, and \textcolor{red}{\sout{that}} \textcolor{blue}{because} we do not
have the beginning and end points of mobile fishing events. Buffer sizes for
fixed and mobile engine was of 200 and 2000 meters, respectively.

@@

\textcolor{red}{\sout{In order to}} \textcolor{blue}{To} characterize the intensity of fishing activities ($FI$), we used
a biomass yield density index. We multiplied the total annual biomass captured
in each grid cell *j*, regardless of species, by the proportion of fishing area
in each grid cell:

@@

The data used to characterize shipping come from the global cumulative impacts
assessment on habitats [@halpern2008a; @halpern2015a] and available on the NCEAS
online data repository [@halpern2015]. We provide a brief summary of data and
methods in this document. For more details, refer to @halpern2015.

@@

The second set of data comes from the Automatic Identification System (AIS),
an initiative launched in 2002 that sought to improve marine safety by providing
mariners with real-time vessel traffic. Through the
International Maritime Organization SOLAS agreement, all vessels of over 300
gross tonnage on international voyages and those carrying passengers are now
required to be equipped with AIS transceivers. These transceivers use Global
Positioning System technology to locate vessels every 10 minutes. The data used
\textcolor{red}{\sout{was}} \textcolor{blue}{were} from November 2010 to December 2011.
Data used come mostly from vessels that move globally (*i.e.* cargo, tanker and
passenger), as they are required to carry AIS transceivers, but also include
data from fishing, high-speed, pleasure and support classes. Shipping intensity
was evaluated as the number of fishing tracks at a 0.1 decimal degrees resolution.
For more details on data and methods used, consult [@walbridge2013].
